---
title: "Exploring BEF-experiment datasets "
author: "Garland Xie"
date: "July 8, 2018"
output: html_document
---

## Load libraries 
```{r libraries}

library(here)
library(readxl)
library(magrittr)
library(tidyverse)

```

Many of the BEF experiments can be loaded via html links. So let's do that.
We'll separate sections by studies then. 

## Cadotte experiment 

Grab the link via DRYAD. (Cadotte et al. 2017. Ecol Lett)

```{r html link 1 }
cadotte_url <-"https://www.datadryad.org/bitstream/handle/10255/dryad.146059/raw_biomass.csv?sequence=1"
```

Load in the Cadotte dataset and let's see the tibble structure. 
Reminder: R recogizes the species richness treatments as integers. Convert into a factor (for now).

```{r importing data, echo=TRUE}
cadotte_df <- read_csv(cadotte_url, 
                          col_types = list(col_character(), 
                                           col_factor(levels = c(1, 2, 3, 4), ordered = TRUE), 
                                           col_double(), 
                                           col_double()
                                           )
                       )

glimpse(cadotte_df)
```

Based on the summary, the necessary data is shown: species richness and biomass.
Note: biomass was sampled using a 0.1 m quadrat in every 2m-square plot 

Next, let's clean up the data a bit and pool monocultures and mixtures appropriately. 

```{r data cleaning, echo = TRUE}
cadotte_clean <- cadotte_df %>% 
  select(SR = Real.rich, biomass = biomass) %>%
  mutate(mixture_type = ifelse(SR > 1, "fine-scale", "coarse-scale")) 

glimpse(cadotte_df)
```

Cool. Let's make a plot per species richness 

```{r plot-cadotte-SR, echo=TRUE}
cadotte_clean %>% 
 ggplot(aes(x = SR, y = biomass, group = SR, col = mixture_type)) +
    geom_jitter(height = 0, width = 0.1, alpha = 0.5) + 
    stat_summary(fun.y = mean,
                 fun.ymin = function(x) mean(x) - sd(x), 
                 fun.ymax = function(x) mean(x) + sd(x), 
                 geom = "pointrange", 
                 colour = "black")  + 
    scale_x_discrete(name = "Species Richness") + 
    scale_colour_discrete(name = "Mixture Type",
                          breaks = c("C", "F"),
                          labels = c("coarse-scale", "fine-scale")) + 
    scale_y_continuous(name = "Biomass (grams per metre-squared)") + 
    theme(panel.background = element_blank())
```

Productivity increases with species richness. Typical result. 
Next, plot by coarse-scale and fine-scale mixture 

```{r plot-cadotte-cofi, echo=TRUE}
cadotte_clean %>% 
 ggplot(aes(x = mixture_type, y = biomass, col = mixture_type)) +
  geom_jitter(height = 0, width = 0.1) + 
  stat_summary(fun.y = mean,
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "pointrange", 
               colour = "black")  + 
  scale_x_discrete(name = "Group-Type") + 
  scale_colour_discrete(name = "Mixture Type",
                        breaks = c("C", "F"),
                        labels = c("coarse-scale", "fine-scale")) + 
  scale_y_continuous(name = "Biomass (grams per metre-squared)") + 
  theme(panel.background = element_blank())
```

## BIODEPTH experiment 

Grab datasets through the ESA archives page. (Spehn et al. 2005. Ecological Monographs)

```{r html links 2}
bd_veg_link <- "http://www.esapubs.org/archive/mono/M075/001/N.vegetation.txt"
bd_design_link <- "http://www.esapubs.org/archive/mono/M075/001/Design.txt" 
```

Let's check out the BIODEPTH biomass dataset 

```{r import BIODEPTH dataset - vegetation}

sr_levels <- c(1, 2, 3, 4, 8, 11, 12, 14, 16, 18, 32)

bd_veg_df <- read_delim(bd_veg_link,
                        delim = "\t", 
                        col_types = list(species.richness = col_factor(levels = sr_levels, 
                                                                       ordered = TRUE)))


glimpse(bd_veg_df)
```

```{r import BIODEPTH dataset - experiment design}
bd_design_df <- read_delim(bd_design_link, delim = "\t")

glimpse(bd_design_df)
```

Might be a good idea to grab the country names for plotting purposes too. 
Note: Maybe I should change the type of join function later on... 

```{r data cleaning}
bd_final_df <- bd_veg_df %>%
  left_join(bd_design_df, by = "plot") %>%
  select(SR = species.richness.x, 
         biomass = mass.g.m2, 
         site = location.y) %>%
  filter(!is.na(biomass)) %>% 
  mutate(mixture_type = ifelse(SR > 1, "F", "C"))

glimpse(bd_final_df)
```

```{r plot-biodepth-cofi-site, fig.height = 10, fig.width = 12}
bd_final_df %>% 
    ggplot(aes(x = mixture_type, y = biomass, col = mixture_type)) +
    geom_jitter(height = 0, width = 0.1, alpha = 0.25) + 
    stat_summary(fun.y = mean,
                 fun.ymin = function(x) mean(x) - sd(x), 
                 fun.ymax = function(x) mean(x) + sd(x), 
                 geom = "pointrange", 
                 colour = "black")  + 
    facet_wrap(.~ site, ncol = 2) + 
    scale_x_discrete(name = "Group-Type") + 
    scale_colour_discrete(name = "Mixture Type",
                        breaks = c("C", "F"),
                        labels = c("coarse-scale", "fine-scale")) + 
    scale_y_continuous(name = "Biomass (grams per metre-squared)") + 
    theme(panel.background = element_blank())
```


## Session info 
```{r sessioninfo, echo = TRUE}
sessionInfo()
```
