---
title: "blesh-eda"
author: "Garland Xie"
date: "04/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(readr)   # for reading comma-delimited text files
library(here)    # for creating relative file paths
library(visdat)  # for visualizing missing data
library(dplyr)   # for manipulating data
library(ggplot2) # for visualizing data
library(forcats) # for manipulating factor variables
```

```{r import}

# raw data
rd_link <- here("data/original", "Blesh_2018_raw_data_species.csv")
raw_df <- read_csv(rd_link)

# metadata
md_link <- here("data/original", "Blesh_2018_metadata.csv")

md_df <- read_csv(md_link)
```

```{r check packging}
# check col data types 
tibble::glimpse(raw_df)

# look for unnecessary rows
head(raw_df, n = 5) 
tail(raw_df, n = 5)
```

```{r check missing data}
vis_dat(raw_df)
vis_miss(raw_df)
```

# Species Info

Sampled species pool:
- CC: Crimson clover (Trifolium incarnatum)
- RC: Medium red clover (Trifolium pratense)
- SW: Spring Wheat (Triticum aestivum)
- WP: Austrian winter pea (Pisum sativum)
- OA: Oat (Avena sativa)
- DR: Daikon radish (Raphanus sativus)
- LN: Lentil (Lens culinaris)
- YM: Yellow mustard (Sinapis alba)
- CV: Chickling vetch (Lathyrus sativus)
- CR: Cereal rye (Secale cereale)

NOTE: clearly, not a natural assemblage (picked crop species)

```{r clean data}
clean_df <- raw_df %>%
  
  # keep appropriate cols
  # FSbiokha = biomass from fall + spring
  select(Farm, Block, Treat, Species, FSbiokha) %>%
  
  # remove weeds and zero biomass data - don't need this
  filter(Species != "WEEDS" &
         FSbiokha != 0) %>%
  
  # assign coarse-het and fine-scale het groups
  # Treat 1: monocultures (WP, CR, SW) - subset of species pool
  mutate(Mosaic = ifelse(Treat > 7, 
                         "coarse-scale", 
                         "fine-scale")) %>%
  
  # manually convert treat ID into species composition
  # easier to interpret BEF trends this way
  mutate(SpComp = case_when( 
                           Treat == 1 ~ "CC + RC + SW", 
                           Treat == 2 ~ "WP + OA + DR", 
                           Treat == 3 ~ "LN + YM + OA", 
                           Treat == 4 ~ "RC + SW", 
                           Treat == 5 ~ "CC + SW", 
                           Treat == 7 ~ "CV + CR", 
                           Treat == 8 ~ "WP", 
                           Treat == 9 ~ "CR", 
                           Treat == 10 ~ "SW")) %>%
  
  # reorder levels by Treat for plotting purposes
  mutate(SpComp = factor(SpComp) %>% fct_reorder(desc(Treat)))
```

```{r summarize data}

# community-level biomass
summ_df <- clean_df %>%
  group_by(Farm, Block, Treat, SpComp, Mosaic) %>%
  summarize(totbio = sum(FSbiokha)) %>%
  ungroup()
```

```{r classic BEF trends}

# species richness vs biomass
summ_df %>%
  ggplot(aes(x = SpComp, y = totbio)) +
  geom_violin() +
  geom_jitter(alpha = 0.1) + 
  stat_summary(fun.y = mean, geom = "point", col = "red") +
  labs(x = "Planted species richness",
       y = "Biomass (kg per ha)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

# species richness vs biomass by BLOCK
# check for variation in block trt
summ_df %>%
  ggplot(aes(x = SpComp, y = totbio)) +
  geom_jitter(alpha = 0.1) + 
  stat_summary(fun.y = mean, geom = "point", col = "red") +
  coord_flip() + 
  facet_wrap(~ Block) + 
  labs(x = "Planted species richness",
       y = "Biomass (kg per ha)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

# conditional plot: species vs richness by FARM
# check for variation in farm trt
summ_df %>%
  ggplot(aes(x = SpComp, y = totbio)) +
  geom_jitter(alpha = 0.1) + 
  stat_summary(fun.y = mean, geom = "point", col = "red") +
  coord_flip() + 
  facet_wrap(~ Farm) + 
  labs(x = "Planted species richness",
       y = "Biomass (kg per ha)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

```{r spat het trends}

# check for any clear diff between biotic het groups 
summ_df %>%
  ggplot(aes(x = factor(Mosaic), y = totbio)) + 
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x),
               fun.ymax = function(x) mean(x) + sd(x),
               geom = "pointrange") +
  labs(x = "Heterogeneity groups", 
       y = "Biomass (kg per ha)")

# check for variation in biotic het groups by BLOCK
summ_df %>%
  ggplot(aes(x = factor(Mosaic), y = totbio)) + 
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x),
               fun.ymax = function(x) mean(x) + sd(x),
               geom = "pointrange") +
  geom_point(alpha = 0.1) + 
  facet_wrap(~ Block) + 
  labs(x = "Heterogeneity groups", 
       y = "Biomass (kg per ha)")

# check for variation in biotic het groups by FARM
summ_df %>%
  ggplot(aes(x = factor(Mosaic), y = totbio)) + 
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x),
               fun.ymax = function(x) mean(x) + sd(x),
               geom = "pointrange") +
  geom_point(alpha = 0.1) + 
  facet_wrap(~ Farm) + 
  labs(x = "Heterogeneity groups", 
       y = "Biomass (kg per ha)")
```

```{r coefficient of variation}
cv_df <- summ_df %>%
  
  # het groups
  group_by(Mosaic) %>%
  
  # get data for meta-analysis  
  summarize(mean  = mean(totbio, na.rm = TRUE) ,
            sd    =  sd(totbio, na.rm = TRUE),
            cv    = sd/mean, 
            n_plots = n(),
            n_trts = length(unique(SpComp))) %>%
  
  mutate(mean = round(mean, 2), 
         sd   = round(sd, 2), 
         cv   = round(cv, 2)) %>%
  
  mutate(years = 1, 
         biome = "Agriculture", 
         study = "Blesh (2018)") %>%
  
  select(study, biome, years, 
         Mosaic, 
         mean, sd, cv, 
         n_plots, n_trts)

cv_df
```

```{r session info}
sessionInfo()
```

